---
title: "ACS data workflow"
author: "Camille Seaberry"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	collapse = TRUE,
	comment = "#>",
	fig.showtext = T
)

```

This is a simple workflow that makes use of functions in `camiller` to analyze and visualize some Census American Community Survey data. The data was downloaded with `tidycensus`, cleaned up a little bit, and is all included in this package.

```{r libraries}
library(tidyverse)
library(camiller)
library(showtext)
```

```{r fonts}
font_add_google("PT Sans", "ptsans")
showtext_auto()
```

The dataset `pov_age` contains estimates and margins of error (MOEs) of residents by age group in different ratio brackets compared to the federal poverty line by town in Greater New Haven.

```{r show_pov_age}
head(pov_age)
```

The function `add_grps` calls the function `make_grps`, which makes it easy to aggregate sums by collapsing multiple subgroups into one larger group. These functions take a list of larger categories, with the indexes of different subgroups in that column. These are easy to figure out by calling `unique` on the column of interest.

```{r uniques}
unique(pov_age$ratio)

unique(pov_age$age)

ratio_grps <- list(
  determined = 1,
  poverty = 2:4,
  low_income = 2:9
)

age_grps <- list(
  young_children = 1,
  children = 1:3,
  seniors = 9:10
)
```

Using `dplyr::group_by(name, ratio)` and then `add_grps` gives aggregates of estimates and MOEs for ratio levels and towns, with the original age groups collapsed into the desired, larger age groups. MOE calculations are done using `tidycensus`. For example:

`c("Under 6 years", "6 to 11 years", "12 to 17 years")`

becomes `children`. The same is then done to collapse ratios. `calc_shares` then calculates shares of residents in each group over the denominator `"determined"`, and, optionally, calculates MOEs for this proportion.

```{r calc_shares}
pov_rates <- pov_age %>%
  mutate_at(vars(age, ratio), function(x) as.factor(x) %>% fct_inorder()) %>%
  group_by(name, ratio) %>%
  add_grps(age_grps, group = age, estimate = estimate, moe = moe) %>%
  group_by(name, age) %>%
  add_grps(ratio_grps, group = ratio, estimate = estimate, moe = moe) %>%
  calc_shares(name, age, group = ratio, denom = "determined", estimate = estimate, moe = moe)
```

```{r pov_rates_out}
head(pov_rates)
```

The function `brk_labels` allows for cleaning up break labels, such as those generated using `cut`. See `?brk_labels` for formatting options. `theme_din` is a clean `ggplot2` theme that works well for dot plots and bar charts.

This legend is unnecessary and not a good idea, just a place to display the output of `brk_labels`.

```{r pov_rate_dots, fig.height=3, fig.width=5}
pov_rates %>%
  filter(ratio == "low_income", age == "children") %>%
  mutate(share = signif(share, digits = 2)) %>%
  mutate(name = as.factor(name) %>% fct_reorder(share, max)) %>%
  arrange(name) %>%
  mutate(brk = classInt::classIntervals(share, n = 3, style = "jenks")$brk %>% cut(share, breaks = ., include.lowest = T)) %>%
  ggplot(aes(x = name, y = share, color = brk)) +
    geom_point(size = 4) +
    coord_flip() +
    scale_color_manual(values = RColorBrewer::brewer.pal(4, "OrRd")[2:4], 
                       labels = function(x) brk_labels(x, format = "percent", mult_by = 100)) +
    theme_din(base_family = "ptsans", base_size = 10, xgrid = T, ygrid = "dotted") +
    scale_y_continuous(labels = scales::percent) +
    labs(x = NULL, y = NULL, color = "Rate", title = "Child low-income rate by town", subtitle = "Greater New Haven towns, 2016", caption = "Source: US Census Bureau 2016 ACS 5-year estimate")
```

The `moe_test` function applies t-tests for differences between two estimates, given their MOEs. This works well for comparing values in one year to those in another, or between related locations or groups. The grunt-work of these calculations is done with `tidycensus` functions. Included in the package is the `pov_trend` tibble for testing out significance testing on data in both 2010 and 2016.

```{r pov_trend}
pov_trend <- pov_age_10_16 %>%
  mutate_at(vars(age, ratio), function(x) as.factor(x) %>% fct_inorder()) %>%
  group_by(name, year, ratio) %>%
  add_grps(age_grps, group = age, estimate = estimate, moe = moe) %>%
  group_by(name, year, age) %>%
  add_grps(ratio_grps, group = ratio, estimate = estimate, moe = moe) %>%
  calc_shares(name, year, age, group = ratio, denom = "determined", estimate = estimate, moe = moe)
```

The output of `moe_test` includes optional intermediary calculations, such as standard errors and Z-scores used for significance testing.

```{r moe_test}
pov_sigs <- pov_trend %>%
  filter(ratio == "low_income") %>%
  select(-estimate, -moe, -ratio) %>%
  mutate(year = paste0("y", year)) %>%
  make_wide(share, sharemoe, group = year) %>%
  moe_test(est1 = y2010_share, moe1 = y2010_sharemoe, est2 = y2016_share, moe2 = y2016_sharemoe, alpha = 0.1)

pov_sigs %>%
  select(name, age, diff:isSig_90) %>%
  head()
```
